---
title: "Food Insecurity at Tract Level"
output: html_document
---

# loag packages
```{r}
library(sf)
library(tidyverse)
library(tmap)
library(tmaptools)
library(tigris)
library(tidycensus)
library(rmapshaper)
library(matrixStats)
library(SpatialAcc)
library(reticulate)
library(dplyr)

library(tidygeocoder)
library(readxl)
```


# load and format Mapping Meal Gap Data
```{r}
mmg2019 <- read_excel("/project/biocomplexity/sdad/projects_data/mc/data_commons/Feeding_America/Map_the_Meal_Gap_Data/MMG2021_2019Data_ToShare.xlsx", sheet = 2)
mmg2019$Food_Insecurity_Rate <- mmg2019$`2019 Food Insecurity Rate`
mmg2019$Child_Food_Insecurity_Rate <- mmg2019$`2019 Child food insecurity rate`

# there was a methodological change in 2018, but I keep the data (not sure how they changed it and if they also changed 2019)
mmg2018 <- read_excel("/project/biocomplexity/sdad/projects_data/mc/data_commons/Feeding_America/Map_the_Meal_Gap_Data/MMG2020_2018Data_ToShare.xlsx", sheet = 1)
mmg2018 <- mmg2018[2:nrow(mmg2018),]
mmg2018$Food_Insecurity_Rate <- mmg2018$...4
mmg2018$Child_Food_Insecurity_Rate <- mmg2018$...13
mmg2018$FIPS <- mmg2018$`Data from MMG 2020 are NOT directly comparable to data from any prior MMG study due to methodological changes made in 2020.`

mmg2017 <- read_excel("/project/biocomplexity/sdad/projects_data/mc/data_commons/Feeding_America/Map_the_Meal_Gap_Data/MMG2019_2017Data_ToShare.xlsx", sheet = 1)
mmg2017$Food_Insecurity_Rate <- mmg2017$`2017 Food Insecurity Rate`
mmg2017$Child_Food_Insecurity_Rate <- mmg2017$`2017 Child food insecurity rate`

mmg2016 <- read_excel("/project/biocomplexity/sdad/projects_data/mc/data_commons/Feeding_America/Map_the_Meal_Gap_Data/MMG2018_2016Data_ToShare.xlsx", sheet = 1)
mmg2016$Food_Insecurity_Rate <- mmg2016$`2016 Food Insecurity Rate`
mmg2016$Child_Food_Insecurity_Rate <- mmg2016$`2016 Child food insecurity rate`

mmg2015 <- read_excel("/project/biocomplexity/sdad/projects_data/mc/data_commons/Feeding_America/Map_the_Meal_Gap_Data/MMG2017_2015Data_ToShare.xlsx", sheet = 1)
mmg2015$Food_Insecurity_Rate <- mmg2015$`2015 Food Insecurity Rate`
mmg2015$Child_Food_Insecurity_Rate <- mmg2015$`2015 Child food insecurity rate`

mmg2014 <- read_excel("/project/biocomplexity/sdad/projects_data/mc/data_commons/Feeding_America/Map_the_Meal_Gap_Data/MMG2016_2014Data_ToShare.xlsx", sheet = 1)
mmg2014$Food_Insecurity_Rate <- mmg2014$`2014 Food Insecurity Rate`
mmg2014$Child_Food_Insecurity_Rate <- mmg2014$`2014 Child food insecurity rate`

mmg2013 <- read_excel("/project/biocomplexity/sdad/projects_data/mc/data_commons/Feeding_America/Map_the_Meal_Gap_Data/MMG2015_2013Data_ToShare.xlsx", sheet = 2)
mmg2013$Food_Insecurity_Rate <- mmg2013$`2013 Food Insecurity Rate`
mmg2013$Child_Food_Insecurity_Rate <- mmg2013$`2013 Child food insecurity rate`

mmg2012 <- read_excel("/project/biocomplexity/sdad/projects_data/mc/data_commons/Feeding_America/Map_the_Meal_Gap_Data/MMG2014_2012Data_ToShare.xlsx", sheet = 2)
mmg2012$Food_Insecurity_Rate <- mmg2012$`2012 Food Insecurity Rate`
mmg2012$Child_Food_Insecurity_Rate <- mmg2012$`2012 Child food insecurity rate`

# we dont have ACS variables needed for these years - not 100%
mmg2011 <- read_excel("/project/biocomplexity/sdad/projects_data/mc/data_commons/Feeding_America/Map_the_Meal_Gap_Data/MMG2013_2011Data_ToShare.xlsx", sheet = 2)
mmg2011$Food_Insecurity_Rate <- mmg2011$`2011 Food Insecurity Rate`
mmg2011$Child_Food_Insecurity_Rate <- mmg2011$`2011 Child Food Insecurity Rate`

mmg2010 <- read_excel("/project/biocomplexity/sdad/projects_data/mc/data_commons/Feeding_America/Map_the_Meal_Gap_Data/MMG2012_2010Data_ToShare.xlsx", sheet = 2)
mmg2010$Food_Insecurity_Rate <- mmg2010$`Food Insecurity Rate`
mmg2010$Child_Food_Insecurity_Rate <- mmg2010$`Child food insecurity rate`

files <- list(mmg2019, mmg2018, mmg2017, mmg2016, mmg2015, mmg2014, mmg2013, mmg2012)
```


# Obtain data to generate model
```{r}
census_api_key("6bc670ff42a8f6682ed80c9ca80ec8c01512088d", overwrite = TRUE)

# start with 2019 and go backward
year <- 2019
all_county_fa_data <- matrix(, nrow = 0, ncol = 13)
for (i in 1:length(files))
{
  my_data_2 <- files[[i]]
    
  # unemployed, non-undergraduate poverty rate, median income, hispanic, black,  HOME OWNERSHIP, disabled
  disabled_variables <- c(m_under_5 = "B18101_004", m_5_17 = "B18101_007", m_18_34 = "B18101_010", m_35_64 = "B18101_013", m_65_74 = "B18101_016", m_above_75 = "B18101_019", f_under_5 = "B18101_023", f_5_17 = "B18101_026", f_18_34 =  "B18101_029", f_35_64 = "B18101_032", f_65_74 = "B18101_035", f_above_75 = "B18101_038", total_ability = "B18101_001")
  
  acs_variables <- get_acs(geography = "county",
                   year = year,
                   variables = c(in_labor_force = "B23025_002",
                                 civilian_unemployed = "B23025_005",
                                 total_income_pop = "B17001_001",
                                 below_poverty_rate = "B17001_002",
                                 undergrads_below_poverty_rate = "B14006_009",
                                 undergrads = "B14001_009",
                                 median_income = "B06011_001",
                                 total_pop = "B01001_001",
                                 hispanic_pop = "B01001I_001",
                                 black_pop = "B01001B_001",
                                 total_home_own = "B07013_001",
                                 home_own = "B07013_002",
                                 disabled_variables),
                   survey = "acs5",
                   output = "wide",
                   geometry = TRUE)
  
  # update acs variables above so that they are in correct format
  acs_variables_update2 <- acs_variables %>%
    mutate(unemp_rate = civilian_unemployedE / in_labor_forceE,
           poverty_rate = (below_poverty_rateE - undergrads_below_poverty_rateE) / (total_income_popE - undergradsE),
           hisp_perc = hispanic_popE / total_popE,
           black_perc = black_popE / total_popE,
           own_perc = home_ownE / total_home_ownE,
           perc_disability = (m_under_5E + m_5_17E + m_18_34E + m_35_64E + m_65_74E + m_above_75E + f_under_5E + f_5_17E + f_18_34E + f_35_64E + f_65_74E + f_above_75E) / total_abilityE) %>%
    rename(median_income = median_incomeE,
           total_pop = total_popE) %>%
    select(GEOID, NAME, unemp_rate, poverty_rate, median_income, hisp_perc, black_perc, own_perc, perc_disability, total_pop)
  
  # change fips to characters and where codes only have 4 characters, add 0 to beginning
  my_data_2$FIPS <- as.character(my_data_2$FIPS)
  idxs <- which(nchar(my_data_2$FIPS) == 4)
  my_data_2$FIPS[idxs] <- paste0(0, my_data_2$FIPS[idxs])
  
  # merge acs data wth food insecurity data from feeding america
  county_fa_data <- merge(st_drop_geometry(acs_variables_update2), my_data_2[, c("Food_Insecurity_Rate", "Child_Food_Insecurity_Rate", "FIPS")], by.x = "GEOID", by.y = "FIPS") %>% mutate(yr = year)
  all_county_fa_data <- rbind(all_county_fa_data, county_fa_data)
  
  year <- year - 1
}
```


# subset for virginia, run glm
```{r}
# subset for 
all_county_fa_data$Food_Insecurity_Rate <- as.numeric(all_county_fa_data$Food_Insecurity_Rate)
all_county_fa_data$state <- substr(all_county_fa_data$GEOID, 1, 2)
va_county_fa_data <- all_county_fa_data[all_county_fa_data$state == "51",]

# I use glm but was unsure if regular lm would be pregered
va_glm3 <- glm(`Food_Insecurity_Rate` ~ GEOID + factor(yr) + unemp_rate + median_income + hisp_perc + black_perc + own_perc + perc_disability, data = va_county_fa_data, weights = total_pop, family = "binomial")

# Kept running into issues with this
# all_glm3 <- glm(`Food_Insecurity_Rate` ~ GEOID + factor(yr) + unemp_rate + median_income + hisp_perc + black_perc + own_perc + perc_disability, data = all_county_fa_data, weights = total_pop, family = "binomial")
```


# get acs information at tract level
```{r}
# unemployed, non-undergraduate poverty rate, median income, hispanic, black,  HOME OWNERSHIP, disabled
tract_acs_variables <- get_acs(geography = "tract",
                               state = 51,
                               year = 2019,
                               variables = c(in_labor_force = "B23025_002",
                                             civilian_unemployed = "B23025_005",
                                             total_income_pop = "B17001_001",
                                             below_poverty_rate = "B17001_002",
                                             undergrads_below_poverty_rate = "B14006_009",
                                             undergrads = "B14001_009",
                                             median_income = "B06011_001",
                                             total_pop = "B01001_001",
                                             hispanic_pop = "B01001I_001",
                                             black_pop = "B01001B_001",
                                             total_home_own = "B07013_001",
                                             home_own = "B07013_002",
                                             disabled_variables),
                               survey = "acs5",
                               output = "wide",
                               geometry = TRUE)

# update acs variables above so that they are in correct format
tract_acs_variables_update2 <- tract_acs_variables %>%
  mutate(unemp_rate = civilian_unemployedE / in_labor_forceE,
         poverty_rate = (below_poverty_rateE - undergrads_below_poverty_rateE) / (total_income_popE - undergradsE),
         hisp_perc = hispanic_popE / total_popE,
         black_perc = black_popE / total_popE,
         own_perc = home_ownE / total_home_ownE,
         perc_disability = (m_under_5E + m_5_17E + m_18_34E + m_35_64E + m_65_74E + m_above_75E + f_under_5E + f_5_17E + f_18_34E + f_35_64E + f_65_74E + f_above_75E) / total_abilityE,
         county = substr(GEOID, 1, 5),
         yr = 2019) %>%
  rename(median_income = median_incomeE,
         tract = GEOID,
         GEOID = county) %>%
  select(tract, GEOID, NAME, unemp_rate, poverty_rate, median_income,
         hisp_perc, black_perc, own_perc, perc_disability, yr)

```


# make predictions for tracts in 2019
```{r}
# histogram to visualize food insecurity estimates
predictions <- predict(va_glm3, tract_acs_variables_update2, type = "response")
hist(predictions)
```


# plot tract predictions for 2019
```{r}
# food insecurity plot at tract level
tract_acs_variables_update3 <- cbind(tract_acs_variables_update2, predictions)
tract_acs_variables_update3 <- tract_acs_variables_update3[!st_is_empty(tract_acs_variables_update3), ]
tract_acs_variables_update3$perc_food_insecure <- tract_acs_variables_update3$predictions * 100
tm_shape(tract_acs_variables_update3, unit = "mi") +
  tm_polygons(col = "perc_food_insecure",
              style = "fixed",
              palette = "Reds",
              border.alpha = 0,
              title = "Food Insecurity (%)",
              breaks = c(0, 5, 10, 20, 30, 65),
              colorNA = "white",
              textNA = "Missing",
              showNA = TRUE) +
  tm_shape(va.ct) +
  tm_polygons(alpha = 0) +
  tm_scale_bar(position = c("left", "top")) +
  tm_layout(main.title = "2019 Food Insecurity Rate",
            main.title.size = 0.95,
            frame = F,
            legend.outside = T,
            legend.outside.position = "right")

tract_acs_variables_update4 <- st_drop_geometry(tract_acs_variables_update3) %>%
  select(tract, NAME, yr, perc_food_insecure) %>%
  rename(geoid = tract,
         region_name = NAME,
         year = yr,
         value = perc_food_insecure) %>%
  mutate(measure = "perc_food_insecure",
         measure_type = "percent",
         measure_units = as.character(NA),
         region_type = "tract") %>%
  relocate("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units")

# tract level estimates for database
tract_acs_variables_update4
```


```{r}
tract_acs_variables_update3
```



# for reference here are the county estimates in 2019
```{r}
my_data_1 <- read_excel("/project/biocomplexity/sdad/projects_data/mc/data_commons/Feeding_America/Map_the_Meal_Gap_Data/MMG2021_2019Data_ToShare.xlsx", sheet = 2)

# xlsx files
va_fa_data <- my_data_1[my_data_1$State == "VA",]
va_fa_data$GEOID <- as.character(va_fa_data$FIPS)

va.co <- get_acs(geography = "county",
                 year = 2019,
                 variables = c(tpop = "B01003_001"),
                 state = "VA",
                 survey = "acs5",
                 output = "wide",
                 geometry = TRUE)

va.co.utm <- st_transform(va.co, crs = "+proj=longlat +datum=WGS84")
va_fa_data_2 <- left_join(va.co.utm, va_fa_data, by = "GEOID")

va_fa_data_2$`2019 Food Insecurity Rate 2` <- va_fa_data_2$`2019 Food Insecurity Rate` * 100

tm_shape(va_fa_data_2, unit = "mi") +
  tm_polygons(col = '2019 Food Insecurity Rate 2',
              style = 'fixed',
              palette = 'Reds',
              border.alpha = 0,
              title = 'Food Insecure (%)',
              breaks = c(0, 5, 10, 20, 30, 65)) +
  tm_shape(co_data_2) +
  tm_polygons(alpha = 0) +
  tm_scale_bar(position = c('left', 'top')) +
  tm_layout(main.title = '2019 Food Insecurity Rate',
            main.title.size = 0.95, frame = F,
            legend.outside = T, legend.outside.position = 'right')

```


# ADDING COUNTY LEVEL FOOD INSECURITY TO THE DATABASE
```{r}
mmg2019 <- read_excel("/project/biocomplexity/sdad/projects_data/mc/data_commons/Feeding_America/Map_the_Meal_Gap_Data/MMG2021_2019Data_ToShare.xlsx", sheet = 2)
mmg2019$Food_Insecurity_Rate <- mmg2019$`2019 Food Insecurity Rate`
mmg2019$Child_Food_Insecurity_Rate <- mmg2019$`2019 Child food insecurity rate`
mmg2019$Cost_Per_Meal <- mmg2019$`2019 Cost Per Meal`
mmg2019$Num_Food_Insecure <- mmg2019$`# of Food Insecure Persons in 2019`
mmg2019$Num_Child_Food_Insecure <- mmg2019$`# of Food Insecure Children in 2019`

# there was a methodological change in 2018, but I keep the data (not sure how they changed it and if they also changed 2019)
mmg2018 <- read_excel("/project/biocomplexity/sdad/projects_data/mc/data_commons/Feeding_America/Map_the_Meal_Gap_Data/MMG2020_2018Data_ToShare.xlsx", sheet = 1)
mmg2018 <- mmg2018[2:nrow(mmg2018),]
mmg2018$Food_Insecurity_Rate <- mmg2018$...4
mmg2018$Child_Food_Insecurity_Rate <- mmg2018$...13
mmg2018$FIPS <- mmg2018$`Data from MMG 2020 are NOT directly comparable to data from any prior MMG study due to methodological changes made in 2020.`
mmg2018$Cost_Per_Meal <- mmg2018$...17
mmg2018$Num_Food_Insecure <- mmg2018$...5
mmg2018$Num_Child_Food_Insecure <- mmg2018$...14

mmg2017 <- read_excel("/project/biocomplexity/sdad/projects_data/mc/data_commons/Feeding_America/Map_the_Meal_Gap_Data/MMG2019_2017Data_ToShare.xlsx", sheet = 1)
mmg2017$Food_Insecurity_Rate <- mmg2017$`2017 Food Insecurity Rate`
mmg2017$Child_Food_Insecurity_Rate <- mmg2017$`2017 Child food insecurity rate`
mmg2017$Cost_Per_Meal <- mmg2017$`2017 Cost Per Meal`
mmg2017$Num_Food_Insecure <- mmg2017$`# of Food Insecure Persons in 2017`
mmg2017$Num_Child_Food_Insecure <- mmg2017$`# of Food Insecure Children in 2017`

mmg2016 <- read_excel("/project/biocomplexity/sdad/projects_data/mc/data_commons/Feeding_America/Map_the_Meal_Gap_Data/MMG2018_2016Data_ToShare.xlsx", sheet = 1)
mmg2016$Food_Insecurity_Rate <- mmg2016$`2016 Food Insecurity Rate`
mmg2016$Child_Food_Insecurity_Rate <- mmg2016$`2016 Child food insecurity rate`
mmg2016$Cost_Per_Meal <- mmg2016$`2016 Cost Per Meal`
mmg2016$Num_Food_Insecure <- mmg2016$`# of Food Insecure Persons in 2016`
mmg2016$Num_Child_Food_Insecure <- mmg2016$`# of Food Insecure Children in 2016`

mmg2015 <- read_excel("/project/biocomplexity/sdad/projects_data/mc/data_commons/Feeding_America/Map_the_Meal_Gap_Data/MMG2017_2015Data_ToShare.xlsx", sheet = 1)
mmg2015$Food_Insecurity_Rate <- mmg2015$`2015 Food Insecurity Rate`
mmg2015$Child_Food_Insecurity_Rate <- mmg2015$`2015 Child food insecurity rate`
mmg2015$Cost_Per_Meal <- mmg2015$`2015 Cost Per Meal`
mmg2015$Num_Food_Insecure <- mmg2015$`# of Food Insecure Persons in 2015`
mmg2015$Num_Child_Food_Insecure <- mmg2015$`# of Food Insecure Children in 2015`

mmg2014 <- read_excel("/project/biocomplexity/sdad/projects_data/mc/data_commons/Feeding_America/Map_the_Meal_Gap_Data/MMG2016_2014Data_ToShare.xlsx", sheet = 1)
mmg2014$Food_Insecurity_Rate <- mmg2014$`2014 Food Insecurity Rate`
mmg2014$Child_Food_Insecurity_Rate <- mmg2014$`2014 Child food insecurity rate`
mmg2014$Cost_Per_Meal <- mmg2014$`2014 Cost Per Meal`
mmg2014$Num_Food_Insecure <- mmg2014$`# of Food Insecure Persons in 2014`
mmg2014$Num_Child_Food_Insecure <- mmg2014$`# of Food Insecure Children in 2014`

mmg2013 <- read_excel("/project/biocomplexity/sdad/projects_data/mc/data_commons/Feeding_America/Map_the_Meal_Gap_Data/MMG2015_2013Data_ToShare.xlsx", sheet = 2)
mmg2013$Food_Insecurity_Rate <- mmg2013$`2013 Food Insecurity Rate`
mmg2013$Child_Food_Insecurity_Rate <- mmg2013$`2013 Child food insecurity rate`
mmg2013$Cost_Per_Meal <- mmg2013$`2013 Cost Per Meal`
mmg2013$Num_Food_Insecure <- mmg2013$`# of Food Insecure Persons in 2013`
mmg2013$Num_Child_Food_Insecure <- mmg2013$`# of Food Insecure Children in 2013`

mmg2012 <- read_excel("/project/biocomplexity/sdad/projects_data/mc/data_commons/Feeding_America/Map_the_Meal_Gap_Data/MMG2014_2012Data_ToShare.xlsx", sheet = 2)
mmg2012$Food_Insecurity_Rate <- mmg2012$`2012 Food Insecurity Rate`
mmg2012$Child_Food_Insecurity_Rate <- mmg2012$`2012 Child food insecurity rate`
mmg2012$Cost_Per_Meal <- mmg2012$`2012 Cost Per Meal`
mmg2012$Num_Food_Insecure <- mmg2012$`# of Food Insecure Persons in 2012`
mmg2012$Num_Child_Food_Insecure <- mmg2012$`# of Food Insecure Children in 2012`

# we dont have ACS variables needed for these years - not 100%
mmg2011 <- read_excel("/project/biocomplexity/sdad/projects_data/mc/data_commons/Feeding_America/Map_the_Meal_Gap_Data/MMG2013_2011Data_ToShare.xlsx", sheet = 2)
mmg2011$Food_Insecurity_Rate <- mmg2011$`2011 Food Insecurity Rate`
mmg2011$Child_Food_Insecurity_Rate <- mmg2011$`2011 Child Food Insecurity Rate`
mmg2011$Cost_Per_Meal <- mmg2011$`Cost Per Meal`
mmg2011$Num_Food_Insecure <- mmg2011$`Number of Food Insecure Persons in 2011`
mmg2011$Num_Child_Food_Insecure <- mmg2011$`Number of Food Insecure Children in 2011`

mmg2010 <- read_excel("/project/biocomplexity/sdad/projects_data/mc/data_commons/Feeding_America/Map_the_Meal_Gap_Data/MMG2012_2010Data_ToShare.xlsx", sheet = 2)
mmg2010$Food_Insecurity_Rate <- mmg2010$`Food Insecurity Rate`
mmg2010$Child_Food_Insecurity_Rate <- mmg2010$`Child food insecurity rate`
mmg2010$Cost_Per_Meal <- mmg2010$`Cost Per Meal`
mmg2010$Num_Food_Insecure <- mmg2010$`Number of Food Insecure Persons`
mmg2010$Num_Child_Food_Insecure <- mmg2010$`Estimated number of food insecure children`

files <- list(mmg2019, mmg2018, mmg2017, mmg2016, mmg2015, mmg2014, mmg2013, mmg2012, mmg2011, mmg2010)
```


```{r}
# start with 2019 and go backward
year <- 2019
all_county_fa_data <- matrix(, nrow = 0, ncol = 5)
for (i in 1:length(files))
{
  my_data_2 <- files[[i]]
    
  # change fips to characters and where codes only have 4 characters, add 0 to beginning
  my_data_2$FIPS <- as.character(my_data_2$FIPS)
  idxs <- which(nchar(my_data_2$FIPS) == 4)
  my_data_2$FIPS[idxs] <- paste0(0, my_data_2$FIPS[idxs])
  
  # merge acs data wth food insecurity data from feeding america
  county_fa_data <- my_data_2[, c("Food_Insecurity_Rate", "Child_Food_Insecurity_Rate", "Cost_Per_Meal", "Num_Food_Insecure", "Num_Child_Food_Insecure", "FIPS")] %>% mutate(yr = year)
  all_county_fa_data <- rbind(all_county_fa_data, county_fa_data)
  
  year <- year - 1
}
```


```{r}
us.co <- get_acs(geography = "county",
                 year = 2019,
                 variables = c(tpop = "B01003_001"),
                 survey = "acs5",
                 output = "wide",
                 geometry = TRUE)

all_county_fa_data_2 <- all_county_fa_data %>%
  mutate(Food_Insecurity_Rate = as.numeric(Food_Insecurity_Rate),
         Child_Food_Insecurity_Rate = as.numeric(Child_Food_Insecurity_Rate),
         Cost_Per_Meal = as.numeric(Cost_Per_Meal)) %>%
  gather(measure, value, c(Food_Insecurity_Rate, Child_Food_Insecurity_Rate, Cost_Per_Meal, Num_Food_Insecure, Num_Child_Food_Insecure)) %>%
  merge(st_drop_geometry(us.co)[, c("GEOID", "NAME")], by.x = "FIPS", by.y = "GEOID") %>%
  rename(geoid = FIPS,
         year = yr,
         region_name = NAME) %>%
  mutate(measure_type = ifelse(measure %in% c("Food_Insecurity_Rate", "Child_Food_Insecurity_Rate"), "percent", ifelse(measure == "Cost_Per_Meal", "cost", "count")),
         measure_units = as.character(NA),
         region_type = "county") %>%
  relocate("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  mutate(value = as.numeric(value)) %>%
  select(-c(tpopE, child_pop))

# source("~/git/VDH/src/helper_functions.R")
# con <- get_db_conn(db_pass = "rsu8zvrsu8zv")
# dc_dbWriteTable(con, "dc_health_behavior_diet", "reg_ct_fa_2010_2019_food_insecurity_and_cost_update3", all_county_fa_data_2)
# dbDisconnect(con)

# con <- get_db_conn(db_pass = "rsu8zvrsu8zv")
# pp <- st_read(con, query = "SELECT * FROM dc_health_behavior_diet.reg_ct_fa_2013_2019_weighted_annual_food_budget_shortfall")
# dbDisconnect(con)
# 
# pp
# 
# virginia_fa_data <- pp[substr(pp$geoid, 1, 2) == "51",]
# ncr_fa_data <- pp %>% filter(geoid %in% c("51013", "51059", "51107", "51510", "51600", "51153", "51683", "51685", "51610", "11001", "24031", "24033", "24017", "24021"))
# 
# con <- get_db_conn(db_pass = "rsu8zvrsu8zv")
# dc_dbWriteTable(con, "dc_health_behavior_diet", "va_ct_fa_2010_2019_food_insecurity_and_cost", va_county_fa_data_db_3)
# dc_dbWriteTable(con, "dc_health_behavior_diet", "ncr_ct_fa_2010_2019_food_insecurity_and_cost", va_county_fa_data_db_3)
# dbDisconnect(con)
```


```{r}
# start with 2019 and go backward
year <- 2019
all_county_fa_data <- matrix(, nrow = 0, ncol = 7)
for (i in 1:length(files))
{
  us.co <- get_acs(geography = "county",
                   year = year,
                   variables = c(tpop = "B01003_001",
                                 male_under_5 = "B01001_003",
                                 male_5_9 = "B01001_004",
                                 male_10_14 = "B01001_005",
                                 male_15_17 = "B01001_006",
                                 female_under_5 = "B01001_027",
                                 female_5_9 = "B01001_028",
                                 female_10_14 = "B01001_029",
                                 female_15_17 = "B01001_030"),
                   survey = "acs5",
                   output = "wide",
                   geometry = TRUE)
  
  us.co2 <- us.co %>%
    mutate(child_pop = male_under_5E + male_5_9E + male_10_14E + male_15_17E + female_under_5E + female_5_9E + female_10_14E + female_15_17E)
           
  my_data_2 <- files[[i]]
    
  # change fips to characters and where codes only have 4 characters, add 0 to beginning
  my_data_2$FIPS <- as.character(my_data_2$FIPS)
  idxs <- which(nchar(my_data_2$FIPS) == 4)
  my_data_2$FIPS[idxs] <- paste0(0, my_data_2$FIPS[idxs])
  
  # merge acs data wth food insecurity data from feeding america
  county_fa_data <- merge(my_data_2[, c("Food_Insecurity_Rate", "Child_Food_Insecurity_Rate", "Cost_Per_Meal", "Num_Food_Insecure", "Num_Child_Food_Insecure", "FIPS")], st_drop_geometry(us.co2)[, c("tpopE", "child_pop", "GEOID")], by.x = "FIPS", by.y = "GEOID") %>% mutate(yr = year)
  all_county_fa_data <- rbind(all_county_fa_data, county_fa_data)
  
  year <- year - 1
}
```

```{r}
health_district <- fread("/project/biocomplexity/sdad/projects_data/vdh/va_county_to_hd.csv")
health_district$county_id <- as.character(health_district$county_id)

va_county_fa_data_db <- all_county_fa_data[substr(all_county_fa_data$FIPS, 1, 2) == "51", ]
va_county_fa_data_db_2 <- va_county_fa_data_db %>%
  merge(health_district, by.x = "FIPS", by.y = "county_id") %>%
  select(-hpr_name) %>%
  mutate(yr = as.character(yr),
         Cost_Per_Meal = as.numeric(Cost_Per_Meal),
         Num_Food_Insecure = as.numeric(Num_Food_Insecure),
         Num_Child_Food_Insecure = as.numeric(Num_Child_Food_Insecure)) %>%
  group_by(health_district, yr) %>%
  summarise(tpop = sum(tpopE),
            Tot_Food_Insecure = sum(Num_Food_Insecure),
            child_pop = sum(child_pop),
            Tot_Child_Food_Insecure = sum(Num_Child_Food_Insecure),
            cost_x_people = sum(Cost_Per_Meal * tpopE))

va_county_fa_data_db_3 <- va_county_fa_data_db_2 %>%
  mutate(Food_Insecurity_Rate = Tot_Food_Insecure / tpop,
         Child_Food_Insecurity_Rate = Tot_Child_Food_Insecure / child_pop,
         Cost_Per_Meal = cost_x_people / tpop) %>%
  select(-c(tpop, child_pop, cost_x_people)) %>%
  rename(Num_Child_Food_Insecure = Tot_Child_Food_Insecure,
         Num_Food_Insecure = Tot_Food_Insecure) %>%
  merge(health_district_geoids, by.x = "health_district", by.y = "region_name") %>%
  gather(measure, value, c(Num_Food_Insecure, Num_Child_Food_Insecure, Food_Insecurity_Rate, Child_Food_Insecurity_Rate, Cost_Per_Meal)) %>%
  mutate(measure_type = ifelse(measure %in% c("Food_Insecurity_Rate", "Child_Food_Insecurity_Rate"), "percent", ifelse(measure == "Cost_Per_Meal", "cost", "count")),
         measure_units = as.character(NA)) %>%
  rename(region_name = health_district,
         year = yr) %>%
  relocate("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units")

con <- get_db_conn(db_pass = "rsu8zvrsu8zv")
dc_dbWriteTable(con, "dc_health_behavior_diet", "va_hd_fa_2010_2019_food_insecurity_and_cost", va_county_fa_data_db_3)
dbDisconnect(con)

head(va_county_fa_data_db_3,1)
head(pp,1)

```


# weighted annual food budget shortfall
```{r}
mmg2019$weighted_budget_shortfall <- mmg2019$`2019 Weighted Annual Food Budget Shortfall`

mmg2018 <- read_excel("/project/biocomplexity/sdad/projects_data/mc/data_commons/Feeding_America/Map_the_Meal_Gap_Data/MMG2020_2018Data_ToShare.xlsx", sheet = 1)
mmg2018 <- mmg2018[2:nrow(mmg2018),]
mmg2018$weighted_budget_shortfall <- mmg2018$...18
mmg2018$FIPS <- mmg2018$`Data from MMG 2020 are NOT directly comparable to data from any prior MMG study due to methodological changes made in 2020.`


mmg2017$weighted_budget_shortfall <- mmg2017$`2017 Weighted Annual Food Budget Shortfall`
mmg2016$weighted_budget_shortfall <- mmg2016$`2016 Weighted Annual Food Budget Shortfall`
mmg2015$weighted_budget_shortfall <- mmg2015$`2015 Weighted Annual Food Budget Shortfall`
mmg2014$weighted_budget_shortfall <- mmg2014$`2014 Weighted Annual Food Budget Shortfall`
mmg2013$weighted_budget_shortfall <- mmg2013$`2013 Weighted Annual Food Budget Shortfall`
files <- list(mmg2019, mmg2018, mmg2017, mmg2016, mmg2015, mmg2014, mmg2013)

# mmg2012$weighted_budget_shortfall <- mmg2012$`2012 Weighted Annual Food Budget Shortfall`
# mmg2011$weighted_budget_shortfall <- mmg2011$`Weighted Annual Food Budget Shortfall`
# mmg2010$weighted_budget_shortfall <- mmg2010$`Weighted Annual Dollars`
# files <- list(mmg2019, mmg2018, mmg2017, mmg2016, mmg2015, mmg2014, mmg2013, mmg2012, mmg2011, mmg2010)


year <- 2019
all_county_fa_data <- matrix(, nrow = 0, ncol = 7)
for (i in 1:length(files))
{
  us.co <- get_acs(geography = "county",
                   year = year,
                   variables = c(tpop = "B01003_001"),
                   survey = "acs5",
                   output = "wide",
                   geometry = TRUE)
  
  my_data_2 <- files[[i]]
    
  print(year)
  # change fips to characters and where codes only have 4 characters, add 0 to beginning
  my_data_2$FIPS <- as.character(my_data_2$FIPS)
  idxs <- which(nchar(my_data_2$FIPS) == 4)
  my_data_2$FIPS[idxs] <- paste0(0, my_data_2$FIPS[idxs])
  
  # merge acs data wth food insecurity data from feeding america
  county_fa_data <- merge(my_data_2[, c("weighted_budget_shortfall", "FIPS")], st_drop_geometry(us.co)[, c("GEOID", "NAME")], by.x = "FIPS", by.y = "GEOID") %>% mutate(yr = year)
  all_county_fa_data <- rbind(all_county_fa_data, county_fa_data)
  
  year <- year - 1
}
```


# WE ALSO WANT WEIGHTED ANNUAL BUDGET SHORTFALL???
```{r}
# merge(st_drop_geometry(us.co[, c("GEOID", "")]), all_county_fa_data, by.x = "GEOID", by.y = "FIPS")

ct.wbs <- all_county_fa_data %>%
  rename(geoid = FIPS,
         region_name = NAME,
         value = weighted_budget_shortfall,
         year = yr) %>%
  mutate(measure = "weighted_budget_shortfall",
         measure_units = as.character(NA),
         measure_type = "dollars",
         region_type = "county") %>%
  relocate("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units")

source("~/git/VDH/src/helper_functions.R")
con <- get_db_conn(db_pass = "rsu8zvrsu8zv")
dc_dbWriteTable(con, "dc_health_behavior_diet", "reg_ct_fa_2011_2019_weighted_annual_food_budget_shortfall", ct.wbs)
dbDisconnect(con)
```

